Using Occupancy Theory and Hierarchical Models to Estimate Otter Population in Southeastern Minnesota \\ {\large Senior Comprehensive Exercise: Carleton College}

### 2010-03-12 Paper.tex
Fixed code and all whitespace problems.

    \subsection{Obtaining convergence on the data}
    Once both the standard hierarchical model and the addition of a CAR prior
    distribution on \(\theta\) to the standard model exhibited convergence on
    our simulated data, we moved on to running both models on the actual data.
    In general, the sparseness of the data gathered in 2004 caused us to
    simplify the model in order to obtain results. Unfortunately, the
    simplification results in estimates of \(\psi\) that do not seem accurate
    and are not directly comparable to estimates of \(\psi\) on 2003's data.

@@

    The 2003 data worked relatively well with the model. The presence of
    numerous
    observations that occur in the second and third days since the most recent
    snow event allows one of the more crucial assumptions of our hierarchical
    model -- that otter tracks laid on each day after a snow event are visible
    throughout the entire observation period and therefore, that it is more
    likely for an observer to detect a track on later days -- to be thoroughly
    tested. Most of the \(\psi\) and \(\theta\) estimates of 2003's data fall
    into ranges where our model is at its best, both with respect to convergence
    and to accurately estimating \(\psi\).

@@

modelUpdate(10000)
samplesSet(c("a", "b", "c", "d", "p[]", "E[]", "psi", "theta"))
#samplesSet(c("a", "b", "p[]", "psi", "theta"))
modelUpdate(10000)
\end{verbatimtab}

>>>>

    \subsection{Obtaining convergence on the data}
    Once both the standard hierarchical model and the addition of a CAR prior
    distribution on \(\theta\) to the standard model exhibited convergence on
    our simulated data, we moved on to running both models on the actual data.
    In general, the sparseness of the data gathered in 2004 caused us to
    simplify the model in order to obtain results. Unfortunately, the
    simplification results in estimates of \(\psi\) that do not seem accurate
    and are not directly comparable to estimates of \(\psi\) on 2003 data.

@@

    The 2003 data worked relatively well with the model. The presence of
    numerous
    observations that occur in the second and third days since the most recent
    snow event allows one of the more crucial assumptions of our hierarchical
    model -- that otter tracks laid on each day after a snow event are visible
    throughout the entire observation period and therefore, that it is more
    likely for an observer to detect a track on later days -- to be thoroughly
    tested. Most of the \(\psi\) and \(\theta\) estimates of 2003 data fall
    into ranges where our model is at its best, both with respect to convergence
    and to accurately estimating \(\psi\).

@@

modelUpdate(10000)
samplesSet(c("a", "b", "c", "d", "p[]", "E[]", "psi", "theta"))
modelUpdate(10000)
\end{verbatimtab}
